Biomass trend maps

### 2019-06-26 analysis/VOCC/biomass-trends.Rmd
make "trawled" time-varying

Species
```{r}
species <- "Quillback Rockfish"
# species <- "Yelloweye Rockfish"
# priors <- TRUE
covariates <- "+trawled+mixed+rocky"

@@

# Choose covariates  
```{r}
covariates <- "+trawled+muddy+rocky+mixed"
covariates <- "+trawled+mixed+rocky"
covariates <- "+muddy+mixed+rocky"
covariates <- "+mixed+rocky"
covariates <- "+trawled+mixed"
```

@@

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

@@

  data <- data %>% gfranges:::scale_survey_predictors()

@@

  if (region == "All synoptic surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
  }

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 300)
  }

@@

    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass, file = paste0("data/", spp, "/mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time1, ".rds"))

@@

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass, file = paste0("data/", spp, "/mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time2, ".rds"))
  } else {
    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

@@

    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
  }
}
```

@@

if (isFALSE(update_spatial_model)) {
  if (region != "All synoptic surveys") {
    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
  } else {
    # Bind together predictions from different models
    try ({
      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-prior-", priors, ".rds"))
      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-prior-", priors, ".rds"))
      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-prior-", priors, ".rds"))
      adult_predictions <- rbind(pred1n3, pred4, pred16) 
      # model_ssid <- unique(adult_predictions$ssid)
      # ssid_string <- paste0(model_ssid, collapse = "n")
    })

@@

    adult_biomass <- readRDS(list.files(
      path = paste0("data/", spp, "/"),
      #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
      pattern = paste0("mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
      all.files = F, full.names = T
    ))

@@

    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions,
      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
    )

    imm_biomass <- readRDS(list.files(
      path = paste0("data/", spp, "/"),
      #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
      pattern = paste0("mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
      all.files = F, full.names = T
    ))
    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions,
      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
    )
  } else {
    rm(adult_predictions)
    rm(imm_predictions)

@@

      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
      nd$year <- as.integer(nd$year)
      predictions <- predict(total_biomass, newdata = nd)
      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
    }
  }
}
```

Bind together predictions from different models
```{r, eval=FALSE}
if (region == "All synoptic surveys") {
  if (isFALSE(update_spatial_model)) {
    try(
      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-mature-biomass-prior-", priors, ".rds")),
      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-mature-biomass-prior-", priors, ".rds")),
      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-mature-biomass-prior-", priors, ".rds")),

@@

    try(
      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-imm-biomass-prior-", priors, ".rds")),
      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-imm-biomass-prior-", priors, ".rds")),
      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-imm-biomass-prior-", priors, ".rds")),

@@

saveRDS(out_mature,
  file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
)

if (!exists("adult_predictions")) {
  out_total <- out_mature
  saveRDS(out_total,
    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
  )
}
```

Optional plot of current total biomass 
```{r eval=FALSE}
if (exists("out_total")) {
  current_biomass <- plot_vocc(out_total,
    fill_col = "var_1_e",
    fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
  current_biomass <- current_biomass +
    ggtitle(paste0("", species, " total biomass"))

  current_biomass
}
```

```{r, error=TRUE}
try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))

if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    # axis_lables = TRUE,
    white_zero = TRUE,
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0("", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))))
  biotrend_ad
} else {
  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))

@@

  out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
  saveRDS(out_imm,
    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
  )
} else {
  rm(out_imm)
}
```

```{r}
try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))

@@

```{r}
png(
  file = paste0("figs/", spp, "/trends-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 11, # The width of the plot in inches
  height = 8
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

@@

png(
  file = paste0("figs/", spp, "/biomass-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 16, # The width of the plot in inches
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

```{r}
png(
  file = paste0("figs/", spp, "/exploratory-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 9.5, # The width of the plot in inches
  height = 12.5
) # The height of the plot in inches

>>>>

Species
```{r}
species <- "Quillback Rockfish"
# species <- "Yelloweye Rockfish"
# priors <- TRUE

@@

# Choose covariates  
```{r}
covariates <- "+muddy+mixed+rocky"
covariates <- "+mixed+rocky"

covariates <- "+trawled+muddy+rocky+mixed"
covariates <- "+trawled+mixed+rocky"
covariates <- "+trawled+mixed"
```

@@

covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

@@

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% gfranges:::scale_survey_predictors()

@@

  # if (region == "All synoptic surveys") {
  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
  # }

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

@@

    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    #saveRDS(adult_biomass, 
    # file = paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, time1, ".rds"))
    saveRDS(adult_biomass, 
      file = paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

@@

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    #saveRDS(imm_biomass, 
    # file = paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, time2, ".rds"))
    saveRDS(imm_biomass, 
     file = paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

  } else {
    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

@@

    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
  }
}
```

```{r eval=FALSE}
k <- length(adult_biomass$model$par)
nll <- adult_biomass$model$objective
AIC <- 2*k+2*nll
AIC #(-5500.438)
```

@@

if (isFALSE(update_spatial_model)) {
  if (region != "All synoptic surveys") {
    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
  } else {
    # Bind together predictions from different models
    try ({
      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-prior-", priors, ".rds"))
      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-prior-", priors, ".rds"))
      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-prior-", priors, ".rds"))
      adult_predictions <- rbind(pred1n3, pred4, pred16) 
      # model_ssid <- unique(adult_predictions$ssid)
      # ssid_string <- paste0(model_ssid, collapse = "n")
    })

@@

    # adult_biomass <- readRDS(list.files(
    #   path = paste0("data/", spp, "/"),
    #   #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
    #   pattern = paste0("mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, "-*"),
    #   all.files = F, full.names = T
    # ))

     adult_biomass <- readRDS(paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

@@

    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions,
      file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
    )

    # imm_biomass <- readRDS(list.files(
    #   path = paste0("data/", spp, "/"),
    #   #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
    #   pattern = paste0("mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, "-*"),
    #   all.files = F, full.names = T
    # ))

    imm_biomass <- readRDS(paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions,
      file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
    )
  } else {
    rm(adult_predictions)
    rm(imm_predictions)

@@

      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
      nd$year <- as.integer(nd$year)
      predictions <- predict(total_biomass, newdata = nd)
      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
    }
  }
}
```

Bind together predictions from different models
```{r, eval=FALSE}
if (region == "All synoptic surveys") {
  if (isFALSE(update_spatial_model)) {
    try(
      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds")),
      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds")),
      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds")),

@@

    try(
      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds")),
      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds")),
      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds")),

@@

saveRDS(out_mature,
  file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
)

if (!exists("adult_predictions")) {
  out_total <- out_mature
  saveRDS(out_total,
    file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
  )
}
```

Attempting to plot of current mature biomass with trawl footprint...
```{r }
# if (exists("out_mature")) {
#   
#   trawl_footprint <- sf::st_read(dsn="data/trawl-footprint",layer="Trawl_footprint")
#   proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
# 
#   trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)
# 
#   
#   current_biomass <- plot_vocc(out_mature,
#     fill_col = "var_1_e",
#     fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
#     raster_alpha = 1,
#     viridis_option = "C",
#     vec_aes = NULL,
#     transform_col = fourth_root_power
#   )
#   current_biomass <- current_biomass +
#   #ggplot(trawl_footprint) + #, aes(x=x/1000, y=y/1000)
#     geom_sf(data= trawl_footprint, aes(geometry=geometry/1000), inherit.aes = FALSE) #+
#     #coord_sf(datum = NULL) 
#   
#   +
#     ggtitle(paste0("", species, " total biomass"))  
# 
#   current_biomass
# }
```

```{r, error=TRUE}
try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))

if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    # axis_lables = TRUE,
    white_zero = TRUE,
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0("", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))))
  biotrend_ad
} else {
  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))

@@

  out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
  saveRDS(out_imm,
    file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
  )
} else {
  rm(out_imm)
}
```

```{r}
try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))

@@

```{r}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 11, # The width of the plot in inches
  height = 8
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

@@

png(
  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 16, # The width of the plot in inches
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

```{r}
png(
  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 9.5, # The width of the plot in inches
  height = 12.5
) # The height of the plot in inches
